\large{Vipassana for Hackers} \\ \Huge{The Proposal} \\ \large\textit{Version 0.1}

### 2024-06-27 the-proposal/the-proposal.tex
add ethics, up to 'timeframes'

@@

\subsubsection{Qualitative}

@@

Data will be collected on a daily basis and themes and narratives regularly collected
with the intention of
describing mental phenomena and the evolution of traits as Vipassana practice
progresses. Monthly or yearly schedules will be decided with a
study supervisor but an example daily schedule
(subject to change) is available in \textit{Appendix 1: Example Daily Schedule}.

\subsubsection{Quantitative}

Measurements performed during the lifelong qualitative study

\subsection{POPULATION AND STUDY SAMPLE}
\subsection{SAMPLE SIZE AND SELECTION OF SAMPLE}
\subsection{SOURCES OF DATA}
\subsection{COLLECTION OF DATA}
\subsection{EXPOSURE ASSESSMENT}
\subsection{DATA MANAGEMENT}
\subsection{DATA ANALYSIS STRATEGIES}

\subsection{Participants}
- people and roles

@@

\subsection{ETHICS AND HUMAN SUBJECTS ISSUES}
\subsection{TIMEFRAMES}

\section{STRENGTHS AND WEAKNESSES OF THE STUDY}
\section{PUBLIC HEALTH SIGNIFICANCE}
\section{BUDGET AND MOTIVATION}
- resources required

>>>>

Outside the scope of Vipassana meditation, the broader topic of
``meditation'' has received a great deal of confused analysis in
recent years.
Many authors will make the mistake of confusing their own subjective
analysis of their personal experiences as somehow objective, or
present it as such.
\cite{ingram2020mastering,https://doi.org/10.1007/s12671-020-01389-4}
It is also an appropriate time to clarify the conclusions of such
confused analysis.

@@

\subsubsection{Qualitative Measures}

@@

Data will be collected on a daily basis and themes and narratives regularly collected
with the intention of
describing mental phenomena and the evolution of traits as Vipassana practice
progresses. Monthly or yearly schedules will be decided with a
study supervisor but an example daily schedule is available in
\textit{Appendix 1: Example Daily Schedule}.

\subsubsection{Quantitative Measures}

Measurements performed during the qualitative study will be used to create a
dataset upon which later inferences will be based.
Measurements taken throughout the study will be subject to equipment availability,
and as such will only be taken based on the future availability of the same
equipment.

In general, quantitative measurements will be taken by additional researchers,
as the study progresses.
This study is, in and of itself, not intended to interpret quantitative measures
of brain activity or otherwise.
Interpretation will be left to researchers familiar with such data, but raw
data will be publicized.

The simplest data available is from off-the-shelf EEG machines, available to
the general public with open-source hardware and open APIs, such as those
sold by BlueBCI \footnote{https://bluebci.com/}, making raw data
easily and permanently accessible to the study.
More advanced EEG machines will require partnership with a university or
research group in the future.

ECG data producing HR and HF-HRV data can be collected from off-the-shelf
systems such as the Zephyr Bioharness 3 \footnote{https://www.zephyranywhere.com/},
previously validated against physiology monitor systems used in laboratories.
\cite{johnstone2012bioharness}

fMRI scans are available at commercial
diagnostic laboratories in India for as little as \$100 USD per scan.
MEG, as a newer technology, will likely require a university partnership.

\subsection{Population and Study Sample}

Initially, only one researcher-subject will be used in this study.

Additional researcher-subjects will be added to the study in the future
if such persons are found capable of dedicating themselves to the task.

\subsection{Collection of Data}

Initial data collection will be done in the form of experiential reports.

\subsection{Data Management}

Data will be the intellectual property of Atapi Research Inc.

Data will be stored and backed up in raw formats produced by diagnostic
equipment under a version controlled data store.

\subsection{Participants}

Initially, the only participant in this study will by me (Steven Deobald).
In the future, additional researchers or research-subjects may be added
to the study.

@@

My willingness to submit myself to this study, for the length of time required,
is based on a number of factors.

First, I have been exclusively practicing Vipassana meditation for over 10 years.
I have taken month-long courses, and will commit myself to taking courses of
one or two months every year for the duration of the study.

Second, I am in the peculiar position of having severe eye damage due to a failed
surgery in 2014.
Although I can still use computers (and will need to do so to write experience
reports), it causes me quite a lot of physical harm to use computers for
extended periods of time.
A severe light sensitivity also caused by the failed surgery means that, peculiarly
enough, my eyes are most comfortable when they are closed in a dark room.

Third, I have been considering this form of research since 2016
--- nearly 8 years at the time of this writing.
I have also seriously considered monastic life, but ordaining as a monk is limiting
in terms of scientific output and there is no monastic order under the Vipassana
lineage yet.

Last, after 10+ years living in India, I am literate in the Devanagari script and
familiar with the Pāli language, which all original Buddhist texts are written in.
While researching meditation directly, it will be important to check findings
against the original teachings to validate my understanding before publishing it.

\subsection{Ethics and Human Subjects Issues}

The ethics of undertaking such a study are personal until and unless another
researcher-subject gets involved.
At that point, those individuals will need to assert for themselves that this
is definitely a process they want to undertake.

That said, there are some personal ethical constraints to address.

The study will leave little room for any other activity.
It is likely that other past-times will need to be nearly eliminated
to pursue the study in earnest.
``Weekends'', as a concept, won't really exist in a study which is intended
to operate continuously.

Other factors which are declared monastic rules, but which seem natural
while practicing deeper meditative states, may come into play.
Monastics are celibate and do not listen to music.
They are also not permitted to run.
While I do intend to become a celibate, eliminate physical exercise,
or stop listening to music, I am cognizant that these are risks of taking
a meditation practice serious to this degree.

\subsubsection{``Getting Paid to Meditate''}

On the subject of \textit{societal} ethics, the topic of
``getting paid to meditate'' has been one I'm acutely aware of, and one
that other commentators on the study have mentioned as well.

If the intention of the study were only to engage in the act of full-time
meditation, I would ordain as a monk.
At present, the system of rules laid out by most monasteries still reflect
those of the Vinaya: a minimum period of 5 years is required to live and
practice independently.
\footnote{https://www.watpahnanachat.org/joining}
Still, if I did not care to publish my investigations into Vipassana
meditation, five years of life is a small price to pay for the
independence to practice continuously.
\cite{nanatusita2014patimokkha}

However, the purpose of financing this study by sponsorship is to pay for
its outputs, not its inputs.
Sponsors who participate are justifiably expecting that the study will
produce readable outcomes, of which they will be the first recipients.

\subsection{Timeframes}

\section{STRENGTHS AND WEAKNESSES OF THE STUDY}

\section{Risks}

* devolves into performance art
* line between science / art is blurry
* forced to take robes
* financially unsustainable

\section{PUBLIC HEALTH SIGNIFICANCE}
\section{BUDGET AND MOTIVATION}
- resources required
